---
title: 'Study 1: Mixed Effects Modeling'
subtitle: Vis 2022 Difference in Uncertainty
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## 1. Import Packages

```{r load-packages}
library(ggplot2)
library(statsr)
library(lme4)
library(sjPlot)
library(dplyr)
library(brmstools)
theme_set(theme_sjplot())
```

## 2. Load Data

```{r load-data}
df <- readr::read_csv("../data/belief_data_prolific_all_exclude.csv")

# refactor and categorize
df$vis_condition <- factor(df$vis_condition, c("uncertainty","scatter","hop"))
levels(df$vis_condition ) <- c("scatter","uncertainty","hop")
```

```{r}
nrow(df)
```

```{r}
# number of uncertainty_difference missing
sum(is.na(df$uncertainty_difference))
```


## 3. Frequentist Mixed Effects Modeling (`lme4`)

```{r lme4-modeling}

# first model is normal response
m1 = lmer(uncertainty_difference ~ vis_condition + (1|user_token), df)

# second model is lognormal response
#m2 = glmer(diffBeliefAbs ~ vis_condition +  (1|user_token), df, family = gaussian(link = "log"))
```

```{r}
a <- plot_model(m1,show.values = TRUE, vline.color = "grey", value.offset = .4, value.size = 3, type="est", show.intercept = TRUE ) +
  scale_y_continuous(breaks=seq(-.75,0.75,.25)) +
  theme(axis.text.y = element_text(size = 8),
        plot.subtitle=element_text(size=11), plot.title = element_text(size = 1)) +
  labs(subtitle = "Uncertainty Difference", title = "") +
  ylim(-0.2, 0.2)

a
```

## 4a. Absolute Belief Difference

### Bayesian Mixed Effects

Let's examine the first regression to estimate the effect on the Uncertainty Difference (`uncertainty_difference`). We'll use the same functional form as model `m`.

```{r eval=FALSE, include=TRUE}
library(brms)

# assume normal response variable
bm <- brms::brm(uncertainty_difference ~ vis_condition + (1|user_token), data = df, backend = "cmdstanr", cores = parallel::detectCores() -1)

save(bm, file = "../models/fit_baseline_diff_uncertainty.rda")
```

```{r}
load("../models/fit_baseline_diff_uncertainty.rda")
```

First let's look at metadata around the model.

```{r}
coef_bm <- coefplot(bm)
coef_bm
```

```{r}
coef_m_df <- a$data %>% rename(Parameter = term) %>% mutate(Parameter = as.character(Parameter))

coef_bm_df <- coef_bm$data
coef_bm_df$Parameter[coef_bm_df$Parameter=="Intercept"] <- "(Intercept)"

joined_models <- inner_join(coef_bm_df, coef_m_df, by = "Parameter")
```


Notice that the coefficients are very similar to Frequentist:

```{r}
joined_models %>%
  rename(Bayesian_Estimate = Estimate, Freq_Estimate = estimate) %>%
  select(Parameter, Bayesian_Estimate, Freq_Estimate) %>%
  mutate(abs_diff = round(abs(Bayesian_Estimate - Freq_Estimate),3)) %>%
  knitr::kable()
```

We see the same for the coefficients standard errors (though they mean slightly different things):

```{r}
joined_models %>%
  rename(Bayesian_Error = Est.Error, Freq_Error = std.error) %>%
  select(Parameter, Bayesian_Error, Freq_Error) %>%
  mutate(abs_diff_error = round(abs(Bayesian_Error - Freq_Error),3)) %>%
  knitr::kable()
```

### Model convergence / posterior predictive check

The convergence stats also look good - Rhat's are at 1 and we have "fuzzy caterpillars". 

```{r}
plot(bm)
```

But remember - convergence doesn't mean great fit. Let's evaluate overfitting with Posterior Predictive Checks. We'll do 10 draws and compare to actual. 

```{r}
pp_check(bm)
```

### Modify response (likelihood) to t-distribution

Let's try instead a distribution likelihood because of the overdispersion (fat tails) we observe in the data.

```{r include=TRUE,eval=FALSE}

bm2 <- brms::brm(uncertainty_difference ~ vis_condition + (1|user_token), data = df, family = student(link = "identity", link_sigma = "log", link_nu = "logm1"), backend = "cmdstanr", cores = parallel::detectCores() - 1)

save(bm2, file = "../models/fit_baseline_diff_uncertainty2.rda")
```

```{r}
load("../models/fit_baseline_diff_uncertainty2.rda")
```

### What are model priors?

```{r}
bm2$prior
```

### What are the coefficients?

```{r}
bm2
```


```{r}
coef_bm2 <- coefplot(bm2)
coef_bm2
```


### Model Comparison

First, let's use leave-one-out (loo) cross-validation. It will also provide estimate to determine point leverage (aka outliers).

```{r}
looNormal <- loo(bm, save_psis = TRUE)
print(looNormal)
```

```{r}
looNormal <- loo(bm, save_psis = TRUE)
print(looNormal)
```

```{r}
looT <- loo(bm2, save_psis = TRUE)
print(looT)
```

When comparing two fitted models, we can estimate the difference in their expected predictive accuracy by the difference in elpd-dloo or elpd-dwaic.

```{r}
loo_compare(looNormal, looT)
```

WAIC criterion

```{r}
waicNormal = waic(bm)
waicT = waic(bm2)
loo_compare(waicNormal, waicT)
```

As a last step, let's do a posterior predictive check:

```{r}
pp_check(bm2) + xlim(-3,3)
```


### Compare Coefficients

As a final check, let's compare the coefficients for the normal Bayesian mixed effects model and the t Distribution Bayesian mixed effects model.

```{r}
coef_bm_df <- coef_bm$data
coef_bm2_df <- coef_bm2$data
coef_bm_df$Parameter[coef_bm_df$Parameter=="Intercept"] <- "(Intercept)"
coef_bm2_df$Parameter[coef_bm2_df$Parameter=="Intercept"] <- "(Intercept)"

joined_models <- inner_join(coef_bm_df, coef_bm2_df, by = "Parameter")
```


Let's examine the coefficient differences with the different likelihoods.

```{r}
un_coef <- joined_models %>%
  rename(Normal_Estimate = Estimate.x, TDist_Estimate = Estimate.y) %>%
  select(Parameter, Normal_Estimate, TDist_Estimate) 

un_error <- joined_models %>%
  rename(Normal_low = `2.5%ile.x`, Normal_high = `97.5%ile.x`,TDist_low = `2.5%ile.y`, TDist_high = `97.5%ile.y`) %>%
  select(Parameter, Normal_low, Normal_high, TDist_low, TDist_high) 

var_order <- c("(Intercept)","vis_conditionuncertainty","vis_conditionhop")

inner_join(un_coef,un_error,by="Parameter") %>%
  tidyr::pivot_longer(-Parameter) %>%
  tidyr::separate(name, c("Model","Estimate"), sep = "_") %>%
  tidyr::pivot_wider(names_from = c("Estimate")) %>%
  mutate(Parameter = factor(Parameter, levels = rev(var_order))) %>%
  mutate(Model = factor(Model, levels = c("Normal", "TDist"))) %>%
  ggplot(aes(x = Parameter, color = Model)) +
  geom_hline(yintercept = 0, alpha = 0.4) +
  geom_point(aes(y = Estimate),  position=position_dodge(.9)) +
  geom_errorbar(aes(ymin = low, ymax = high),  position=position_dodge(.9)) +
  theme(legend.position = c(0.2,0.85),) +
  labs(title = "Uncertainty Difference", subtitle = "Vary by Response Distribution") +
  scale_color_manual(values = c("TDist" = "red",
                                "Normal"="black")) +  
  coord_flip()
```

We see the same for the coefficients standard errors:

```{r}
joined_models %>%
  rename(Normal_Error = Est.Error.x, Lognormal_Error = Est.Error.y) %>%
  select(Parameter, Normal_Error, Lognormal_Error) %>%
  mutate(Diff_Error = round(Normal_Error - Lognormal_Error,3)) %>%
  knitr::kable()
```


## Candidate models

For model selection, we will consider additional models.

* `uncertainty_difference ~ vis_condition + (1|user_token) + true_correlation * vis_condition`

```{r include=TRUE,eval=FALSE}
# https://discourse.mc-stan.org/t/smooth-spline-modeling-with-brm/6364
bm3 <- brms::brm(uncertainty_difference ~ vis_condition + (1|user_token) + true_correlation * vis_condition, data = df, family = student(link = "identity", link_sigma = "log", link_nu = "logm1"),, backend = "cmdstanr", cores = parallel::detectCores() - 1)

save(bm3, file = "../models/fit_baseline_diff_uncertainty3.rda")
```

* `uncertainty_difference ~ vis_condition + (1|user_token) + pre_belief_distance * vis_condition`

```{r include=TRUE,eval=FALSE}
bm4 <- brms::brm(uncertainty_difference ~ vis_condition + (1|user_token) + pre_belief_distance * vis_condition, data = df, family = student(link = "identity", link_sigma = "log", link_nu = "logm1"), backend = "cmdstanr", cores = parallel::detectCores() - 1)

save(bm4, file = "../models/fit_baseline_diff_uncertainty4.rda")
```

* `uncertainty_difference ~ vis_condition + (1|user_token) + pre_belief_distance * vis_condition + true_correlation * vis_condition`

```{r include=TRUE,eval=FALSE}
bm5 <- brms::brm(uncertainty_difference ~ vis_condition + (1|user_token) + pre_belief_distance * vis_condition + true_correlation * vis_condition, data = df, family = student(link = "identity", link_sigma = "log", link_nu = "logm1"),, backend = "cmdstanr", cores = parallel::detectCores() - 1)

save(bm5, file = "../models/fit_baseline_diff_uncertainty5.rda")
```

* `uncertainty_difference ~ vis_condition + (1|user_token) + pre_attitude * vis_condition`

```{r include=TRUE,eval=FALSE}
bm6 <- brms::brm(uncertainty_difference ~ vis_condition + (1|user_token) + pre_attitude_strength * vis_condition, data = df, family = student(link = "identity", link_sigma = "log", link_nu = "logm1"),, backend = "cmdstanr", cores = parallel::detectCores() - 1)

save(bm6, file = "../models/fit_baseline_diff_uncertainty6.rda")
```

```{r}
load("../models/fit_baseline_diff_uncertainty3.rda")
load("../models/fit_baseline_diff_uncertainty4.rda")
load("../models/fit_baseline_diff_uncertainty5.rda")
load("../models/fit_baseline_diff_uncertainty6.rda")
```

```{r}
waic3 = waic(bm3)
waic4 = waic(bm4)
waic5 = waic(bm5)
waic6 = waic(bm6)
```

```{r}
loo_compare(waicNormal, waicT, waic3, waic4, waic5, waic6)
```


```{r}
pp_check(bm6) + xlim(-3,3)
```

```{r}
bm5
```

```{r}
coef_bm5 <- coefplot(bm5)
coef_bm5
```

```{r}
bm6
```


```{r}
coef_bm6 <- coefplot(bm6)
coef_bm6
```

## Posterior Predictives

```{r}
library(bayesplot)

# bm2

mcmc_areas(
  bm2,
  pars = c("b_vis_conditionuncertainty","b_vis_conditionhop"),
  prob = 0.8, # 80% intervals
  prob_outer = 0.99, # 99%
  point_est = "mean"
)
```

```{r}
# bm5

mcmc_areas(
  bm5,
  pars = c("b_vis_conditionuncertainty","b_vis_conditionhop"),
  prob = 0.8, # 80% intervals
  prob_outer = 0.99, # 99%
  point_est = "mean"
)
```


